fGFPCA dynamic prediction (In-sample)

### 2023-01-09 f-GFPCA.Rmd
try out-of-sample mutlidimensional integral

Let say on the binned grid, at each interval s, new sample has $n_s$ observations, among which $m_s$ are successes. $m_s = \sum_{j=1}^{n_S} I(Y_{sj}=1)$, j is the index for observation in a specific interval.

@@

$$\begin{aligned}
p(\boldsymbol{Y}|\boldsymbol{\xi}) &= \prod_{s=1}^{m}p(\boldsymbol{Y}_t|\boldsymbol{\xi})
\end{aligned}$$
<!-- We can utilize the fitted series of local GLMMs, instead of refitting them. Since $\boldsymbol{Y}_{new}$ is a binary function, we can get its (incomplete) latent function value $\eta^s$ with Maximum Likelihood Estimation. Specifically, we need to estimate its (incomplete) random intercept function $b^s$. -->

@@

1. Write out the formula for the BLUPs in the local GLMMs. Since we have a random intercept only model this should not be too complicated. (this is for out of sample prediction only)
2. Approximate these values numerically via numeric integration (like we did witht he AUC paper) and/or explicit laplace approximation 
3. Evaluate performance of dynamic predictions (both using "in sample" data -- those individuals whose BLUPs we can extract directly from the local glmm fits and out of sample data)
4. Write an R package (see Julia's fastGFPCA and Andrew's fcr pacakge as inspiration) which has functions for performing dynamic prediction using our framework
5. Implement a brief simulation study showing both that the approach works

>>>>

Let say on the binned grid, at each interval s, new sample has $n_s$ observations, among which $h_s$ are successes. $h_s = \sum_{j=1}^{n_S} I(Y_{sj}=1)$, j is the index for observation in a specific interval.

@@

$$\begin{aligned}
p(\boldsymbol{Y}|\boldsymbol{\xi}) &= \prod_{s=1}^{m}p(\boldsymbol{Y}_s|\boldsymbol{\xi}) \\
& = \prod_{s=1}^{m} \binom{h_s}{n_s}p_s^{h_s}(1-p_s)^{n_s-h_s}\\
logit(p_s) & = \hat{\eta}_s = \hat{f}_{0}(s)+\phi(s)\boldsymbol{\xi}
\end{aligned}$$

Please note that here, we are assuming $\boldsymbol{Y}_s$ is independent from each other given fPCA score. That is, all correlation across time within the same subject is introduced by underlying eigenfunctions.

If we write this down for a single score:

$$E(\xi_k|\boldsymbol{Y})=\frac{\int \xi_k \int p(\boldsymbol{Y}|\boldsymbol{\xi})p(\boldsymbol{\xi})d\boldsymbol{\xi_{(-k)}}d\xi_k}{\int_{\boldsymbol{\xi}}p(\boldsymbol{Y}|\boldsymbol{\xi})p(\boldsymbol{\xi})d\boldsymbol{\xi}}$$

# Problems: 
Numeric integration too slow (adaptIntegrate multidimensional)! Maybe try  MLE?

<!-- We can utilize the fitted series of local GLMMs, instead of refitting them. Since $\boldsymbol{Y}_{new}$ is a binary function, we can get its (incomplete) latent function value $\eta^s$ with Maximum Likelihood Estimation. Specifically, we need to estimate its (incomplete) random intercept function $b^s$. -->

@@

1. Write out the formula for the BLUPs in the local GLMMs. Since we have a random intercept only model this should not be too complicated. (this is for out of sample prediction only)

2. Approximate these values numerically via numeric integration (like we did witht he AUC paper) and/or explicit laplace approximation 
3. Evaluate performance of dynamic predictions (both using "in sample" data -- those individuals whose BLUPs we can extract directly from the local glmm fits and out of sample data)
4. Write an R package (see Julia's fastGFPCA and Andrew's fcr pacakge as inspiration) which has functions for performing dynamic prediction using our framework
5. Implement a brief simulation study showing both that the approach works
